ff
``` {r}
library(tidyverse)
library(janitor)
```

``` {r}
setwd('/Users/joellevtov/Library/CloudStorage/OneDrive-JoelLev-Tov/Journalism/delaware_discipline')
```

``` {r}
delaware_discipline <- read_csv("source_files/delaware_student_discipline_raw.csv") |> 
    clean_names() |> 
    select(!geography) #the geography column reads either "All Students" or, for redacted rows, "NA" - so this will ignore the column
delaware_discipline$sub_group <- str_replace(delaware_discipline$sub_group, regex("Asian", ignore_case = TRUE), "Asian American") 
delaware_discipline$category <- str_replace(delaware_discipline$category, regex("Out-of-School Suspension, No CDAP Placement", ignore_case = TRUE), "Out-of-School Suspension without CDAP Placement")

# The raw data has two different phrases that mean the same thing - "Out-of-School Suspension, No CDAP Placement" and "Out-of-School Suspension without CDAP Placement". For clarity, I replaced the latter with the former.
delaware_pop <- read_csv("source_files/Student_Enrollment_20240216.csv") |> 
    clean_names()
delaware_pop$grade <- str_replace(delaware_pop$grade, regex("Twelfth", ignore_case=TRUE), "12th Grade")
delaware_pop$sub_group <- str_replace(delaware_pop$sub_group, regex("Asian", ignore_case = TRUE), "Asian American")
```

Non-function tests
``` {r}
subgroup_test <- delaware_discipline |> 
    filter(str_detect(sub_group, regex("homeless", ignore_case = TRUE)))
print(subgroup_test)

subgroup_test <- delaware_discipline |> 
    filter(str_detect(sub_group, "All Students")) |> 
    select(sub_group)
print(subgroup_test)

non_in_subgroup_query_test <- delaware_discipline |> 
    filter(str_detect(sub_group, regex("homeless", ignore_case = TRUE))) |> 
    filter(!str_detect(sub_group, regex("non", ignore_case = TRUE))) |> 
    select(sub_group)
print(non_in_subgroup_query_test)
```

Function test
This function will allow us to search inside the sub_group category, for example for homeless students. This one is a bit different than the other columns because the entries often have a 'non' inside, i.e. non-homeless or non-SWD. That presents a bit of a challenge because I have made the code search for anything that contains the query - think of a Cntl+F search in a PDF - it will show you anything that contains the characters you're looking for. 
``` {r}
# WORK IN PROGRESS
subgroupdata <- function(subgroup_query, race_name, pl[discipline_df) {
    # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
     looking_for_non <- function(subgroup_query, discipline_df) {
        filtered <- discipline_df |> 
            filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
        return(filtered)
        }

    if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
        filtered <- discipline_df |> 
            filter(sub_group==race_name)
    }
    # This elif block only runs if there is NO 'non' in the query and the query isn't "all" or null 
    else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
        filtered <- discipline_df |>
            filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
            filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
    # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
    } 
    else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
        # Add hyphen after "non"
        subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
        print(subgroup_query)
        filtered <- looking_for_non(subgroup_query, discipline_df)
    # The else block runs in the case that there is a non already with a hyphen
} 
else {
    filtered <- looking_for_non(subgroup_query, discipline_df)
}
    return(filtered)
} 
results <- subgroupdata(subgroup_query="all", delaware_discipline) |> select(sub_group)
head(results)
```


LINE 146

I think this function could be simplified a lot. It runs through the identical process for each type of discipline, it just takes the median of different columns. So wouldn't it make more sense for it to take the median of (provided_column) or something like that?
yes, claude says, something like this would work:calculate_median <- function(df, column_name) 
    median(na.omit(df[[column_name]]))
``` {r}
process_race_data <- function(pop_df, discipline_df, dis_code, race_name=NULL, subgroup_query=NULL, grade_query=NULL, gender_query=NULL, disc_type=NULL) {

    do_analysis <- function(provided_df, dis_code, race_name=NULL, subgroup_query=NULL, grade_query=NULL, gender_query=NULL, disc_type=NULL) {        
        # Dis_code filter
        if (is.null(dis_code) || tolower(dis_code) == "all") {
            # If it's null or 'all', the person wants to look at discipline statewide. The function could do nothing, but leaving the field blank would not achieve that - it would be an average of statewide and of each individual district
            filtered <- provided_df |> 
            filter(district_code==0)
        } else {
            filter(district_code== dis_code)
        }
        
        # Race_name filter
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |> 
                filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name)

        # Sub_group filter
        subgroupdata <- function(subgroup_query, provided_df) {
            looking_for_non <- function(subgroup_query, provided_df) {
                filtered <- provided_df |> #might have to be provided_df actually, not filtered
                    filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
                return(filtered)
            } 
            if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
                filtered <- provided_df
            } 
            else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
                filtered <- provided_df |>
                    filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                    filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
            # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
            } 
            else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
                # Add hyphen after "non"
                subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
                print(subgroup_query)
                filtered <- looking_for_non(subgroup_query, provided_df)
            # The else block runs in the case that there is a non already with a hyphen
            } else {
                filtered <- looking_for_non(subgroup_query, provided_df)
            }
                return(filtered)
            } 
            filtered <- filtered |> 
                subgroupdata(subgroup_query, provided_df)
        }
        
        # Grade_query filter
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- provided_df |>  
                filter(grade == "All Students")
        } else {
            filtered <- provided_df |>  
            filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }

        # Gender_query filter
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- provided_df |>  
            filter(gender == "All Students")
        } else {
            filtered <- provided_df |>  
            filter(gender == gender_query)
        }
    
        # Calculating median
        median_students <- median(na.omit(filtered$provided_column))
        return(median_students)
    }
    provided_column <- 
    # this function calculates the length of a discipline measure
    length_of_discipline <- function(discipline_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {
        median_discipline_length <- median(na.omit(filtered$avg_duration), na.rm=TRUE)
        return(median_discipline_length)
    }
    # this function calculates the number of students in a given circumstance, e.g. the number of Black students in fourth grade
    find_pop <- function(pop_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {}

    num_incidents <- find_num_incidents(discipline_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)
    num_disc_students <- find_num_disc_students(discipline_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)
    pop <- find_pop(pop_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)
    length_discipline <- length_of_discipline(discipline_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)

    incidents_per_race <- data.frame (
        students_enrolled = as.numeric(pop),
        num_disc_students = as.numeric(num_disc_students),
        num_incidents = as.numeric(num_incidents),
        length_in_days = as.numeric(length_discipline))

    incidents_per_race <- incidents_per_race |> 
        mutate(per_capita = as.numeric(num_incidents) / as.numeric(pop))    
}
```

``` {r}
subgroupdata <- function(subgroup_query, provided_df) {
        # Instead of defining this function each and every damn time, it makes more sense to have it search the provided dataframe
        # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
        looking_for_non <- function(subgroup_query, provided_df) {
            filtered <- provided_df |> 
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
            return(filtered)
        } 
    # this function calculates the average number of incidents for each racial category
    find_num_incidents <- function(discipline_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {
        # Start with filtering district_code
        filtered <- discipline_df |> 
            filter(district_code == dis_code) 
    
        # Apply filter for race_name. Defaults to "All Students" if "all" or NULL
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |>  
                filter(race == "All Students")
        } 
        else {
            filtered <- filtered |> 
                filter(race == race_name)
        }
        # Applying a filter for sub_group. It's mostly useless, honestly, because for most groups - even at the state level - the data on race + population is redacted (i.e., the number of e.g. homeless students who are Asian)
        if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
            filtered <- discipline_df |> 
                filter(sub_group=="All students")
        } 
        else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
            filtered <- discipline_df |>
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
        # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
        } 
        else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
            # Add hyphen after "non"
            subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
            print(subgroup_query)
            filtered <- looking_for_non(subgroup_query, discipline_df)
        } 
        # The else block runs in the case that there is a non already with a hyphen
        else {
            filtered <- looking_for_non(subgroup_query, discipline_df)
        }
            return(filtered)
        } 
        # Apply filter for grade_query. Uses str_detect for partial matches in case race name isn't spelled out like database lists it. Default to "All Students" if "all" or NULL
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- filtered |>  
            filter(grade == "All Students")
        } else {
            filtered <- filtered |>  
            filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }
        # Apply filter for gender_query. Defaults to "All Students" if "all" or NULL
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |> 
            filter(gender == "All Students")
        } else {
            filtered <- filtered |> 
            filter(gender == gender_query)
        }
        # Apply filter for disc_type if provided and not "all". No "All Students" default for discipline type because there is no equivalent
        if (!is.null(disc_type) && tolower(disc_type) != "all") {
            filtered <- filtered |> 
            filter(category == disc_type)
        } 
        
        median_incidents <- median(na.omit(filtered$incidents))
    }    

    # this function calculates the average number of students disciplined for the provided filters
    find_num_disc_students <- function(discipline_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {
        filtered <- discipline_df |> 
            filter(district_code == dis_code) 
    
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |>  
            filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name) 
        } 
        subgroupdata <- function(subgroup_query, discipline_df) {
        # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
        looking_for_non <- function(subgroup_query, discipline_df) {
            filtered <- discipline_df |> 
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
            return(filtered)
        } 
        if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
            filtered <- discipline_df |> 
                filter(sub_group==race_name)
        } 
        else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
            filtered <- discipline_df |>
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
        # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
        } 
        else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
            # Add hyphen after "non"
            subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
            print(subgroup_query)
            filtered <- looking_for_non(subgroup_query, discipline_df)
        # The else block runs in the case that there is a non already with a hyphen
        } else {
            filtered <- looking_for_non(subgroup_query, discipline_df)
        }
            return(filtered)
        }        
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- filtered |>  
            filter(grade == "All Students")
        } else {
            filtered <- filtered |>  
            filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }        
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |> 
            filter(gender == "All Students")
        } else {
            filtered <- filtered |> 
            filter(gender == gender_query)
        }    
        if (!is.null(disc_type) && tolower(disc_type) != "all") {
            filtered <- filtered |> 
            filter(category == disc_type)
        } 
        median_disc_students <- median(na.omit(filtered$students))
        return(median_disc_students)
```