ff
``` {r}
library(tidyverse)
library(janitor)
```

``` {r}
setwd('/Users/joellevtov/Library/CloudStorage/OneDrive-JoelLev-Tov/Journalism/delaware_discipline')
```

``` {r}
delaware_discipline <- read_csv("source_files/delaware_student_discipline_raw.csv") |> 
    clean_names() |> 
    select(!geography) |>  #the geography column reads either "All Students" or, for redacted rows, "NA" - so this will ignore the column 
    select(!sub_group) # This column presents so many problems but is just a summary of all the other columns so we wil lremove it
delaware_discipline$category <- str_replace(delaware_discipline$category, regex("Out-of-School Suspension, No CDAP Placement", ignore_case = TRUE), "Out-of-School Suspension without CDAP Placement")

# The raw data has two different phrases that mean the same thing - "Out-of-School Suspension, No CDAP Placement" and "Out-of-School Suspension without CDAP Placement". For clarity, I replaced the latter with the former.
delaware_pop <- read_csv("source_files/Student_Enrollment_20240216.csv") |> 
    clean_names() |> 
    select(!sub_group)
delaware_pop$grade <- str_replace(delaware_pop$grade, regex("Twelfth", ignore_case=TRUE), "12th Grade")
```

Non-function tests
``` {r}
```

``` {r}
native_am_test <- process_race_data(pop_df=delaware_pop, discipline_df = delaware_discipline, dis_code=0, race_name=name_native_am, schoolyear="2016,2017")
print(native_am_test)

hisp_test <- process_race_data(pop_df = delaware_pop, discipline_df = delaware_discipline, dis_code = 0, race_name = name_hisp) 
print(hisp_test)

```

LINE 146

++++++ NEW +++++++++
I think this function could be simplified a lot. It runs through the identical process for each type of discipline, it just takes the median of different columns. So wouldn't it make more sense for it to take the median of (provided_column) or something like that?
yes, claude says, something like this would work:calculate_median <- function(df, column_name) 
    median(na.omit(df[[column_name]]))
``` {r}
process_race_data <- function(pop_df, discipline_df, dis_code=NULL, race_name=NULL, subgroup=NULL, grade_query=NULL, gender_query=NULL, disc_type=NULL, schoolyear=NULL) { 
    do_analysis <- function(provided_df, dis_code=NULL, race_name=NULL, subgroup=NULL, grade_query=NULL, gender_query=NULL, disc_type=NULL, provided_column, schoolyear=NULL) {        
        
    message("\nStarting do_analysis with column: ", provided_column)
    message("Initial rows: ", nrow(provided_df))
    message("Unique special_demo values before filtering: ", paste(unique(provided_df$special_demo), collapse=", "))
        # Dis_code filter
        if (is.null(dis_code) || tolower(dis_code) == "all") {
            # If it's null or 'all', the person wants to look at discipline statewide. The function could do nothing, but leaving the field blank would not achieve that - it would be an average of statewide and of each individual district
            filtered <- provided_df |> 
                filter(district_code==0)
        } else {
            filtered <- provided_df |> 
            filter(district_code== dis_code)
        }
        message("After district filter: ", nrow(filtered))
        message("Unique special_demo values after race filter: ", paste(unique(filtered$special_demo), collapse=", "))
        
        message("schoolyear value: ", schoolyear)
        message("schoolyear class: ", class(schoolyear))
        # School year filter
            if (is.null(schoolyear) || str_detect(schoolyear, regex("all", ignore_case = TRUE))) {
                over_1_year <- "yes"
            } else if (nchar(schoolyear) > 4) {
                # Split by comma and trim whitespace
                years <- schoolyear |>
                    str_split(",") |>
                    unlist() |>
                    str_trim()
                message(years, collapse=", ")
                
            # Check if all years are valid
            valid_years <- c("2015","2016","2017","2018","2019","2020","2022","2023")
            if (!all(years %in% valid_years)) {
                invalid_years <- years[!years %in% valid_years]
                stop("Invalid year(s): ", paste(invalid_years, collapse=", "))
            }
            
            filtered <- filtered |> 
                filter(school_year %in% years)
            over_1_year <- "yes"
        } else {
            filtered <- filtered |> 
                filter(school_year==schoolyear)
            over_1_year <- "no"
        }
        message("After school year filter: ", nrow(filtered))
        message("Unique special_demo values after gender_query: ", paste(unique(filtered$special_demo), collapse=", "))

        # Race_name filter
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |> 
                filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name)
        }
        message("After race filter: ", nrow(filtered))
        message("Unique special_demo values after race filter: ", paste(unique(filtered$special_demo), collapse=", "))
    
        # Grade_query filter
        if (is.null(grade_query) || str_detect(grade_query, regex("all", ignore_case = TRUE))) {
            filtered <- filtered |>  
                filter(grade == "All Students")
        } else {
            filtered <- filtered |>  
                filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }
        message("After grade filter: ", nrow(filtered))
        message("Unique special_demo values after grade_query: ", paste(unique(filtered$special_demo), collapse=", "))

        # Gender_query filter
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |>  
            filter(gender == "All Students")
        } else {
            filtered <- filtered |>  
                filter(gender == gender_query)
        }
        message("After gender filter: ", nrow(filtered))
        message("Unique special_demo values after gender_query: ", paste(unique(filtered$special_demo), collapse=", "))

        # Apply filter for disc_type if provided and not "all". No "All Students" default for discipline type because there is no equivalent
        if (!is.null(disc_type) && tolower(disc_type) != "all") {
            filtered <- filtered |> 
                filter(category == disc_type)
        } 
        message("After discipline type filter: ", nrow(filtered))

        # looking_for_non function
        looking_for_non <- function(subgroup, provided_df) {
            filtered <- filtered |> 
                filter(str_detect(special_demo, regex(subgroup, ignore_case = TRUE)))
            return(filtered)
        }
        
        # special_demo filter
        message("Unique special_demo values: ", paste(unique(filtered$special_demo), collapse=", "))
        if (is.null(subgroup) || str_detect(subgroup, regex("all", ignore_case=TRUE))){
            # If a user is interested in one of the categories that isn't goverened by special_demo, like Black female first graders, this will ensure that the averages are accurate 
            # by filtering out 'All students' - that way, the number of students and the number of incidents isn't doubled
            if (is.null(schoolyear) || str_detect(schoolyear, regex("all", ignore_case = TRUE))) {
                filtered <- filtered |> 
                    filter(!special_demo==regex("All Students", ignore_case = TRUE))
            } else {
            }
        } else if (str_detect(subgroup, regex("non", ignore_case = TRUE), negate = TRUE)) {
            # If there is a value provided like 'homeless' - filters out anything with a 'non'
            filtered <- filtered |>
                filter(str_detect(special_demo, regex(subgroup, ignore_case = TRUE)))  |>
                filter(!str_detect(special_demo, regex("non", ignore_case = TRUE)))
        } else if (str_detect(subgroup, regex("non(?!-)", ignore_case = TRUE))) {
            # If there is a 'non' and there's no hyphen, add hyphen to make it 'non-'
            subgroup <- str_replace(subgroup, regex("non(?!-)", ignore_case = TRUE), "non-") 
            print(subgroup)
            filtered <- looking_for_non(subgroup, provided_df)
        } else {
            # If the 'non' already has a hyphen
            filtered <- looking_for_non(subgroup, provided_df)
        }
        message("After special demo filter: ", nrow(filtered))
        
    # Calculating median for each aspect
    provided_column <- as.character(provided_column)
    # message("Final number of rows before median: ", nrow(filtered))
    # message("Values in column ", provided_column, ": ", paste(head(filtered[[provided_column]]), collapse=", "))
    if (over_1_year=="yes") {
        result <- filtered |>
            group_by(school_year) |>  # First group by school year
            summarize(year_median = median(na.omit(.data[[provided_column]]))) |>  # Get median for each year
            ungroup() |>  # Ungroup to operate on all years
            summarize(final_median = median(year_median)) |>  # Get median of the yearly medians
            pull(final_median)  # Extract the final value
    } else {
        result <- median(na.omit(filtered[[provided_column]]))
    }

    return(result)
    }

    num_disc_students <- do_analysis(provided_df = discipline_df, dis_code=dis_code, race_name=race_name, subgroup=subgroup, grade_query=grade_query, gender_query=gender_query, disc_type=disc_type, provided_column = "students", schoolyear)
    length_in_days <- do_analysis(provided_df = discipline_df, dis_code=dis_code, race_name=race_name, subgroup=subgroup, grade_query=grade_query, gender_query=gender_query, disc_type=disc_type, provided_column = "avg_duration", schoolyear)
    students_enrolled <- do_analysis(provided_df = pop_df, dis_code=dis_code, race_name=race_name, subgroup=subgroup, grade_query=grade_query, gender_query=gender_query, provided_column = "students", schoolyear, disc_type=NULL)
    median_num_incidents <- do_analysis(provided_df = discipline_df, dis_code=dis_code, race_name=race_name, subgroup=subgroup, grade_query=grade_query, gender_query=gender_query, disc_type=disc_type, provided_column = "incidents", schoolyear)


    incidents_per_race <- data.frame (
        students_enrolled = as.numeric(students_enrolled),
        num_disc_students = as.numeric(num_disc_students),
        num_incidents = as.numeric(median_num_incidents),
        length_in_days = as.numeric(length_in_days))

    incidents_per_race <- incidents_per_race |> 
        mutate(per_capita = as.numeric(num_incidents) / as.numeric(students_enrolled))    
}
```


++++ OLD ++++++ 
+++++ FOR COMPARISON +++++++
``` {r}
process_race_data_old <- function(pop_df, discipline_df, dis_code, race_name=NULL, subgroup_query=NULL, grade_query=NULL, gender_query=NULL, disc_type=NULL) {
    message("\nStarting find_num_incidents")
    message("Initial rows: ", nrow(discipline_df))
    # Function definitions
    # this function calculates the average number of incidents for each racial category
    find_num_incidents <- function(discipline_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {
        # Start with filtering district_code
        filtered <- discipline_df |> 
            filter(district_code == dis_code) 
        message("After district filter: ", nrow(filtered))
    
        # Apply filter for race_name. Defaults to "All Students" if "all" or NULL
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |>  
            filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name)
        }
        message("After race filter: ", nrow(filtered))
        
        # Apply filter for grade_query. Uses str_detect for partial matches in case race name isn't spelled out like database lists it. Default to "All Students" if "all" or NULL
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- filtered |>  
            filter(grade == "All Students")
        } else {
            filtered <- filtered |>  
            filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }
        message("After grade filter: ", nrow(filtered))


        # Apply filter for gender_query. Defaults to "All Students" if "all" or NULL
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |> 
            filter(gender == "All Students")
        } else {
            filtered <- filtered |> 
            filter(gender == gender_query)
        }
        message("After gender filter: ", nrow(filtered))

        # Apply filter for disc_type if provided and not "all". No "All Students" default for discipline type because there is no equivalent
        if (!is.null(disc_type) && tolower(disc_type) != "all") {
            filtered <- filtered |> 
            filter(category == disc_type)
        } 
        message("After discipline type filter: ", nrow(filtered))

        # Applying a filter for sub_group. It's mostly useless, honestly, because for most groups - even at the state level - the data on race + population is redacted (i.e., the number of e.g. homeless students who are Asian)
        subgroupdata <- function(subgroup_query, discipline_df) {
        # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
        looking_for_non <- function(subgroup_query, discipline_df) {
            filtered <- filtered |> 
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
            return(filtered)
        } 

        if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
            filtered <- discipline_df |> 
                filter(sub_group==race_name)
        } else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
            filtered <- discipline_df |>
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
        # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
        } else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
            # Add hyphen after "non"
            subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
            print(subgroup_query)
            filtered <- looking_for_non(subgroup_query, discipline_df)
        # The else block runs in the case that there is a non already with a hyphen
        } else {
            filtered <- looking_for_non(subgroup_query, discipline_df)
        }
            return(filtered)
        } 
        
        median_incidents <- median(na.omit(filtered$incidents))
    }    

    # this function calculates the average number of students disciplined for the provided filters
    find_num_disc_students <- function(discipline_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {
        filtered <- discipline_df |> 
            filter(district_code == dis_code) 
    
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |>  
            filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name) 
        } 
        subgroupdata <- function(subgroup_query, discipline_df) {
        # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
        looking_for_non <- function(subgroup_query, discipline_df) {
            filtered <- discipline_df |> 
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
            return(filtered)
        } 
        if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
            filtered <- discipline_df |> 
                filter(sub_group==race_name)
        } 
        else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
            filtered <- discipline_df |>
                filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
        # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
        } 
        else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
            # Add hyphen after "non"
            subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
            print(subgroup_query)
            filtered <- looking_for_non(subgroup_query, discipline_df)
        # The else block runs in the case that there is a non already with a hyphen
        } else {
            filtered <- looking_for_non(subgroup_query, discipline_df)
        }
            return(filtered)
        }        
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- filtered |>  
            filter(grade == "All Students")
        } else {
            filtered <- filtered |>  
            filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }        
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |> 
            filter(gender == "All Students")
        } else {
            filtered <- filtered |> 
            filter(gender == gender_query)
        }    
        if (!is.null(disc_type) && tolower(disc_type) != "all") {
            filtered <- filtered |> 
            filter(category == disc_type)
        } 
        median_disc_students <- median(na.omit(filtered$students))
        return(median_disc_students)
    }    

    # this function calculates the number of students in a given circumstance, e.g. the number of Black students in fourth grade
    find_pop <- function(pop_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) { 
        filtered <- pop_df |> 
            filter(district_code== dis_code)
        
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |> 
                filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name)
        }

        subgroupdata <- function(subgroup_query, pop_df) {
            # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
            looking_for_non <- function(subgroup_query, pop_df) {
                filtered <- pop_df |> 
                    filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
                return(filtered)
            if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
                filtered <- pop_df
            } else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
                filtered <- pop_df |>
                    filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                    filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
            # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
            } else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
                # Add hyphen after "non"
                subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
                print(subgroup_query)
                filtered <- looking_for_non(subgroup_query, pop_df)
            # The else block runs in the case that there is a non already with a hyphen
            } else {
                filtered <- looking_for_non(subgroup_query, pop_df)
            }
            return(filtered)
            } 
        }

        # grade_query filter
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- filtered |>  
            filter(grade == "All Students")
        } else {
            filtered <- filtered |>  
            filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }

        #gender_query filter
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |>  
            filter(gender == "All Students")
        } else {
            filtered <- filtered |>  
            filter(gender == gender_query)
        }

        median_students <- median(na.omit(filtered$students))
        return(median_students)
    }

    # this function calculates the length of a discipline measure
    length_of_discipline <- function(discipline_df, dis_code, race_name = NULL, subgroup_query=NULL, grade_query = NULL, gender_query = NULL, disc_type = NULL) {
        filtered <- discipline_df |> 
            filter(district_code == dis_code) 
        
        # As above, will default to all students if there's no value provided
        if (is.null(race_name) || tolower(race_name) == "all") {
            filtered <- filtered |>  
            filter(race == "All Students")
        } else {
            filtered <- filtered |> 
            filter(race == race_name) 
        }
        subgroupdata <- function(subgroup_query, discipline_df) {
        # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
            looking_for_non <- function(subgroup_query, discipline_df) {
                filtered <- discipline_df |> 
                    filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
                return(filtered)
            } 
            if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
                filtered <- discipline_df |> 
                    filter(sub_group==race_name)
            } else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
                filtered <- discipline_df |>
                    filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
                    filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
            # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
            } else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
                # Add hyphen after "non"
                subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
                print(subgroup_query)
                filtered <- looking_for_non(subgroup_query, discipline_df)
            # The else block runs in the case that there is a non already with a hyphen
            } else {
                filtered <- looking_for_non(subgroup_query, discipline_df)
            }
                return(filtered)
        } 
        # The same is true for the grade
        if (is.null(grade_query) || tolower(grade_query) == "all") {
            filtered <- filtered |>  
            filter(grade == "All Students")
        } else {
            filtered <- filtered |>  filter(str_detect(grade, regex(grade_query, ignore_case = TRUE)))
        }
        # The same is true for the gender (note that there's only male and female in the data)
        if (is.null(gender_query) || tolower(gender_query) == "all") {
            filtered <- filtered |> 
            filter(gender == "All Students")
        } else {
            filtered <- filtered |> 
            filter(gender == gender_query)
        }
        # This filters for what type of discipline (i.e. out of school suspension, in-school suspension, etc.)
        if (!is.null(disc_type) && tolower(disc_type) != "all") {
            filtered <- filtered |> 
            filter(category == disc_type)
        } else {
            stop("That discipline type doesn't exist.")
        }
        median_discipline_length <- median(na.omit(filtered$avg_duration), na.rm=TRUE)
        return(median_discipline_length)
    }

    num_incidents <- find_num_incidents(discipline_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)
    num_disc_students <- find_num_disc_students(discipline_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)
    pop <- find_pop(pop_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)
    length_discipline <- length_of_discipline(discipline_df, dis_code, race_name, subgroup_query, grade_query, gender_query, disc_type)

    incidents_per_race <- data.frame (
        students_enrolled = as.numeric(pop),
        num_disc_students = as.numeric(num_disc_students),
        num_incidents = as.numeric(num_incidents),
        length_in_days = as.numeric(length_discipline))

    incidents_per_race <- incidents_per_race |> 
        mutate(per_capita = as.numeric(num_incidents) / as.numeric(pop))    
}
```


``` {r}
colnames(delaware_discipline)
testing_function <- delaware_discipline |> 
    filter(district_code==0)  |> 
    filter(school_year=="2016" | school_year=="2017") |> 
    print() |> 
    distinct(school_year) |> 
    print()
```

``` {r}
runschoolyear <- function(schoolyear, df){
    if (is.null(schoolyear) || str_detect(schoolyear, regex("all", ignore_case = TRUE))) {
        over_1_year <- "yes"
    } else if (nchar(schoolyear) > 4) {
        # Split by comma and trim whitespace
        years <- schoolyear |>
            str_split(",") |>
            unlist() |>
            str_trim()
        message(paste(years, collapse=", "))
            
        # Check if all years are valid
        valid_years <- c("2015","2016","2017","2018","2019","2020","2022","2023")
        if (!all(years %in% valid_years)) {
            invalid_years <- years[!years %in% valid_years]
            stop("Invalid year(s): ", paste(invalid_years, collapse=", "))
        }
            
        filtered <- df |> 
            filter(school_year %in% years)
        over_1_year <- "yes"
    } else {
        filtered <- df |> 
            filter(school_year==schoolyear)
        over_1_year <- "no"
    }
    return(filtered)
}
print(runschoolyear("2016, 2017", delaware_discipline) |>  distinct(school_year))
```

``` {r}
result <- delaware_discipline |> 
    filter(district_code==0) |> 
    filter(race==name_hisp) |> 
    filter(grade=="All Students") |> 
    filter(gender=="All Students") |> 
    filter(!special_demo==regex("All Students", ignore_case = TRUE)) |>  # Add this
    group_by(school_year) |>
    print() 
    #summarize(year_median = median(na.omit(incidents))) |>
    #ungroup() |>
    #summarize(final_median = median(year_median))
#print(result)
```
hisp_incidents <- process_race_data(pop_df = delaware_pop, discipline_df = delaware_discipline, dis_code = 0, race_name = name_hisp) # providing no arguments for grade_query, gender_query, disc_type, and subgroup as they are automatically set to 'null'
 [1] "school_year"    "district_code"  "district"       "school_code"   
 [5] "organization"   "race"           "gender"         "grade"         
 [9] "special_demo"   "sub_group"      "category"       "rowstatus"     
[13] "students"       "enrollment"     "pct_enrollment" "incidents"     
[17] "avg_duration"