ff
``` {r}
library(tidyverse)
library(janitor)
```

``` {r}
setwd('/Users/joellevtov/Library/CloudStorage/OneDrive-JoelLev-Tov/Journalism/delaware_discipline')
```

``` {r}
delaware_discipline <- read_csv("source_files/delaware_student_discipline_raw.csv") |> 
    clean_names() |> 
    select(!geography) #the geography column reads either "All Students" or, for redacted rows, "NA" - so this will ignore the column
delaware_discipline$sub_group <- str_replace(delaware_discipline$sub_group, regex("Asian", ignore_case = TRUE), "Asian American") 
delaware_discipline$category <- str_replace(delaware_discipline$category, regex("Out-of-School Suspension, No CDAP Placement", ignore_case = TRUE), "Out-of-School Suspension without CDAP Placement")

# The raw data has two different phrases that mean the same thing - "Out-of-School Suspension, No CDAP Placement" and "Out-of-School Suspension without CDAP Placement". For clarity, I replaced the latter with the former.
delaware_pop <- read_csv("source_files/Student_Enrollment_20240216.csv") |> 
    clean_names()
delaware_pop$grade <- str_replace(delaware_pop$grade, regex("Twelfth", ignore_case=TRUE), "12th Grade")
delaware_pop$sub_group <- str_replace(delaware_pop$sub_group, regex("Asian", ignore_case = TRUE), "Asian American")
```

Non-function tests
``` {r}
subgroup_test <- delaware_discipline |> 
    filter(str_detect(sub_group, regex("homeless", ignore_case = TRUE)))
print(subgroup_test)

subgroup_test <- delaware_discipline |> 
    filter(str_detect(sub_group, "All Students")) |> 
    select(sub_group)
print(subgroup_test)

non_in_subgroup_query_test <- delaware_discipline |> 
    filter(str_detect(sub_group, regex("homeless", ignore_case = TRUE))) |> 
    filter(!str_detect(sub_group, regex("non", ignore_case = TRUE))) |> 
    select(sub_group)
print(non_in_subgroup_query_test)
```

Function test
This function will allow us to search inside the sub_group category, for example for homeless students. This one is a bit different than the other columns because the entries often have a 'non' inside, i.e. non-homeless or non-SWD. That presents a bit of a challenge because I have made the code search for anything that contains the query - think of a Cntl+F search in a PDF - it will show you anything that contains the characters you're looking for. 
``` {r}
# WORK IN PROGRESS
subgroupdata <- function(subgroup_query, discipline_df) {
    # The below function will handle any queries that have a 'non' inside of them, e.g. 'non-homeless'
     looking_for_non <- function(subgroup_query, discipline_df) {
        filtered <- discipline_df |> 
            filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))
        return(filtered)
        }

    if (is.null(subgroup_query) || str_detect(subgroup_query, regex("all", ignore_case=TRUE))){
        filtered <- discipline_df
    }
    # This elif block only runs if there is NO 'non' in the query and the query isn't "all" or null 
    else if (str_detect(subgroup_query, regex("non", ignore_case = TRUE), negate = TRUE)) { 
        filtered <- discipline_df |>
            filter(str_detect(sub_group, regex(subgroup_query, ignore_case = TRUE)))  |>
            filter(!str_detect(sub_group, regex("non", ignore_case = TRUE)))
    # This else if block detects whether the query contains 'non' without a hyphen - if it does, it will add a hyphen after 'non' and run the looking_for_non function 
    } 
    else if (str_detect(subgroup_query, regex("non(?!-)", ignore_case = TRUE))) {
        # Add hyphen after "non"
        subgroup_query <- str_replace(subgroup_query, regex("non(?!-)", ignore_case = TRUE), "non-") 
        print(subgroup_query)
        filtered <- looking_for_non(subgroup_query, discipline_df)
    # The else block runs in the case that there is a non already with a hyphen
} 
else {
    filtered <- looking_for_non(subgroup_query, discipline_df)
}
    return(filtered)
} 
results <- subgroupdata(subgroup_query="all", delaware_discipline) |> select(sub_group)
head(results)
```


LINE 146

how would i write r code that looks for if the query entered no hyphen AND contains "non". if this is true, i need it to add a hyphen after the non-. here is what i have so far:
``` {r}

```